### 2014-10-26 thesis/method/preprocess.tex
Misc changes to preprocessing and selection

\section{Normalisation}
Features represent different pieces of information about a training example.
Therefore, it is not surprising that not all features will have the same range
of values: a patient can have an \texttt{age} feature with a value of 20, but
their \texttt{heart rate} feature cannot possibly be 20. Learning algorithms
such as $k$-nearest neighbour (Chapter \ref{chap:classification})
are sensitive to the scales of feature values -- features with a larger range
of values will dominate a distance calculation. Therefore, we
\textit{normalise} all numeric (that is, continuous-valued)
feature values into the $[-1,1]$ range so that they are all expressed on the
same scale. Typically, values are normalised in the $[0,1]$ or the $[-1,1]$
ranges.
The formula to do this, also discussed in Chapter \ref{chap:classification},
is:
\begin{equation*}
a_i = 2\left[\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right] - 1
\end{equation*}
where we have used $s=2$ and $t=-1$ to transform values into the $[-1,1]$
range. This transformation was applied to each value of all numeric features,
and was performed after all cleaning steps.

>>>>

At the end of the cleaning process, the data was converted into the
Attribute-Relation File Format (ARFF)
\footnote{\url{http://www.cs.waikato.ac.nz/ml/weka/arff.html}} for use in
WEKA, a suite of machine learning algorithms implemented in Java
\cite{Hall2009}. Each feature was carefully checked to ensure that their
type was specified correctly -- that is, the nominal attributes were
distinguished from the numeric.

\section{Normalisation}
Features represent different pieces of information about a training example.
Therefore, it is not surprising that not all features will have the same range
of values: a patient can have an \texttt{age} feature with a value of 20, but
their \texttt{heart rate} feature cannot possibly be 20. Learning algorithms
such as $k$-nearest neighbour (Chapter \ref{chap:classification})
are sensitive to the scales of feature values -- features with a larger range
of values will dominate a distance calculation. Therefore, we
\textit{normalise} all numeric (that is, continuous-valued)
feature values into the $[-1,1]$ range so that they are all expressed on the
same scale. Typically, values are normalised in the $[0,1]$ or the $[-1,1]$
ranges.
The formula to do this, also discussed in Chapter \ref{chap:classification},
is:
\begin{equation*}
a_i = 2\left[\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right] - 1
\end{equation*}
where we have used $s=2$ and $t=-1$ to transform values into the $[-1,1]$
range. This transformation was applied to each value of all numeric features
using WEKA, and was performed after all cleaning steps described above.
